Competing Bandits

### 2018-08-03 paper/competing_bandits.tex
updates to paper

\subsection{Bandit Priors}
We study the implications of our model via simulation. For computational tractability we focus our attention on Bernoulli bandits with beta priors. We consider the following bandit ``priors" from which the bandit instances faced in simulation are drawn:
\begin{enumerate}
\item ``HeavyTail": We took the mean rewards to be randomly drawn from Beta($\alpha=0.6,\beta=0.6$). With this distribution it was likely to have arms that were at the extremes (close to 1 and close to 0) but also some of the arms with intermediate value means.
\item Needle-in-haystack High - $K-1$ arms with mean 0.50, 1 arm with mean 0.70. 
\end{enumerate}

@@

We look at the performance of each of our learning algorithms in isolation, meaning that we simply run the learning algorithm for $T$ steps on a realization of the prior instances that we consider and record the realized rewards and the arm that is pulled. In our model the reputation score is an input to the decision rule of the agents and thus this statistic is an important property of the learning algorithm selected by the arm. Figure \ref{prelim_means} displays the mean average reputation score of each of the algorithms across the different simulations.

@@

As expected we see that according to the mean reputation plots $TS > DEG > DG$ for sufficiently large values of $t$. In the competition game we have defined, however, another relevant statistic is the relative comparison of reputation values across algorithms. Figure \ref{prelim_comparison} displays the proportion of rounds where one algorithm has at least as high of a reputation as another. Interestingly we observe that, on the Heavy Tail prior, $DG > DEG$ according to this metric even though $DEG > DG$ according to the mean reputation. The intuition for this is simple. $DG$ will either find the best arm or not. If it does then since $DEG$ engages in non-adaptive exploration $DG$ will, with $\epsilon$ probability, pick an arm uniformly at random which potentially hurts its reputation if the selected arm is not the best arm. $DEG$ eventually converges on the best arm with a sufficiently large time horizon, but $DG$ does not necessarily do so and so when $DG$ does not find the best arm it may potentially have substantially worse reputation which drags down its mean reputation. However, if it is able to find the best arm in sufficiently many simulations then it may have slightly higher reputation than $DEG$ in those simulations and thus do better according to the relative reputation proportion but, due to the simulations where it does not find the best arm it may then have worse mean reputation.

[Would be nice to get at what is the driving factor behind this. Right now, we simply observe it.]

@@

In this section we present simulation results of the model discussed in Section \ref{S:2} where $X = 0$ so that both firms enter at the same time. We want to understand what algorithms the firms are incentivized to commit to in the competition game and so we take the algorithms as given and study the consequences of playing them in the competition game. For this experiment we fix the $HardMax$ agent decision rule. Under this decision rule we expect that one firm will end up taking the market in most of the simulations and so when reasoning about the decision rule we track the effective end of game.
[Reword this]

@@

Tables \ref{hm_nih} and \ref{hm_ht} display the results. Under Heavy Tail we see that for small warm start of $k = 20$, $DG$ gives the firm more of the market than $TS$ and $DEG$. As we increase the warm start to $k= 100, 200$ we observe that $TS$ does better than $DG$ and $DEG$. Note that, in particular for the simulations involving $TS$, the effective end of game had a median of 0. This means that the result of the simulation was \textit{completely determined by the warm start period}. Thus, when reasoning about what algorithm would win in the competition game under the HardMax decision rule we need to think about the performance of the algorithm on $k$ samples, which may be small. For low warm start, it is thus not necessarily the case that $TS$ is the best algorithm. The intuition for this is that $TS$ needs to engage in pure exploration early on in order to identify the best arm. A consequence of exploration in our model is that it may lead to a reputational cost relative to the greedy algorithm since it involves experimenting with potentially sub-optimal arms relative to the current belief of the best arm. However, a sufficiently large warm start allows $TS$ to incur this reputational cost without the informational gain.

Another interesting result is that $DG$ still does better than $DEG$ even with a large warm start under the Heavy Tail prior. In general, we observe that even under other agent response functions besides HardMax that $DG$ does better than $DEG$. Thus, even though the mean reputation plot in Figure \ref{prelim_means} shows that $DEG$ does better than $DG$ in terms of average mean reputation, this does not lead to $DEG$ doing better under competition. Rather, the relative reputation plot shown in Figure \ref{relative_rep} is more predictive in this case of what algorithm wins under competition.

Looking at the results for Needle In a Haystack we observe that $TS$ always does better than $DEG, DG$. T

[TODO: is there something more interesting / general that we can discuss here besides this? Why does TS win? Learning is easy? Something completely unrelated to learning hardness?]

@@

[This section merits much more explanation]

>>>>

\subsection{Bandit Priors}
We study the implications of our model via simulation. For computational tractability we focus our attention on Bernoulli bandits with beta priors. We consider the following bandit ``priors" from which the bandit instances faced in simulation are drawn:
\begin{enumerate}
\item ``HeavyTail": We took the mean rewards to be randomly drawn from Beta($\alpha=0.6,\beta=0.6$). With this distribution it was likely to have arms that were at the extremes (close to 1 and close to 0) but also some of the arms with intermediate value means. [We should rename this prior instance]
\item Needle-in-haystack High - $K-1$ arms with mean 0.50, 1 arm with mean 0.70. 
\end{enumerate}

[Need some discussion of what makes these priors at all ``representative" - why can we just focus on them?]

@@

We look at the performance of each of our learning algorithms in isolation, meaning that we simply run the learning algorithm for $T$ steps on a realization of the prior instances that we consider and record the realized rewards and the arm that is pulled. In our simulations the realizations of each of the arms were pre-drawn and the same across all algorithms and so the differences in algorithm performance come from the difference in the arms selected by the algorithms and not from differences in realizations. When looking at the results of each of the learning algorithms in isolation we want to confirm that, for the prior instances that we consider, ``better" algorithms win according to ``absolute" measures of performance (i.e. looking at mean reputation). Additionally, we want to understand what statistics we can look at in isolation to understand what wins under competition.

In our model the reputation score is an input to the decision rule of the agents and thus it is important to understand how this score differs across algorithms when run outside of competition. Figure \ref{prelim_means} displays the mean average reputation score of each of the algorithms across the different simulations.

@@

As expected we see that according to the mean reputation plots $TS > DEG > DG$ for sufficiently large values of $t$. While looking at the mean reputation is a natural statistic to consider when attempting to understand what algorithm wins under competition, we consider an additional statistic when attempting to understand what wins under competition. In the competition game the decision rule of the agents compares the relative reputation of the two firms so that if the reputation of firm $A$ is higher than that of firm $B$, the agents will (be more likely to) select firm $A$. Thus, another natural statistic to look at is to look at the two algorithms in isolation and consider the fraction of simulations where, for a fixed time $t$, the reputation of firm $A$ is at least as high as that of firm $B$. Figure \ref{prelim_comparison} displays the proportion of rounds where one algorithm has at least as high of a reputation as another.

Interestingly we observe that, on the Heavy Tail prior, $DG > DEG$ according to this metric even though $DEG > DG$ according to the mean reputation. The intuition for this is simple. $DG$ will either find the best arm or not. If it does then since $DEG$ engages in non-adaptive exploration $DG$ will, with $\epsilon$ probability, pick an arm uniformly at random which potentially hurts its reputation if the selected arm is not the best arm. $DEG$ eventually converges on the best arm with a sufficiently large time horizon, but $DG$ does not necessarily do so and so when $DG$ does not find the best arm it may potentially have substantially worse reputation which drags down its mean reputation. However, if it is able to find the best arm in sufficiently many simulations then it may have slightly higher reputation than $DEG$ in those simulations and thus do better according to the relative reputation proportion but, due to the simulations where it does not find the best arm it may then have worse mean reputation. In the next section we move to considering the competition game and revisit if this statistic is more useful for predicting who wins in competition.

@@

In this section we present simulation results of the model discussed in Section \ref{S:2} where $X = 0$ so that both firms enter at the same time. We want to understand what algorithms the firms are incentivized to commit to in the competition game but the payoffs associated with each algorithm are too difficult to compete analytically and thus, we begin by simply imposing that the algorithm played by the firm is taken as given and study the consequences of playing them in the competition game. Additionally, for this experiment we present results for the $HardMax$ agent decision rule and postpone discussion of alternative decision rules. Besides keeping track of market share, we also will keep track of the effective end of game in order to better understand what happens in the game.

@@

\begin{figure}
\label{rel_rep_ht_fine}
\caption{Relative Reputation - Heavy Tail}
\includegraphics[scale=0.3]{figures/ts_dg_heavy_fine}
\caption*{TODO: Reduce the range of the x-axis of this so that you can understand when TS starts to do better than DG}
\end{figure}

Tables \ref{hm_nih} and \ref{hm_ht} display the results. Under Heavy Tail we see that for small warm start of $k = 20$, $DG$ gives the firm more of the market than $TS$ and $DEG$. As we increase the warm start to $k = 100, 200$ we observe that $TS$ does better than $DG$ and $DEG$. Note that, in particular for the simulations involving $TS$, the effective end of game had a median of 0. This means that the result of the simulation was \textit{completely determined by the warm start period}. Thus, when reasoning about what algorithm would win in the competition game under the HardMax decision rule we need to think about the performance of the algorithm on $k$ samples instead of $T$ samples, which may be small. For low warm start, it is thus not necessarily the case that $TS$ is the best algorithm. However, we see that under the Needle In Haystack prior $TS$ is always the best algorithm across all warm start values.

[Missing here is a discussion of the fact that with HardMax the market gets taken by one firm completely (in most simulations) as well as why this is the case.]

First, we note that the relative reputation plots were better at predicting the result of the competition game than the ``absolute" mean reputation plots. For the Needle In Haystack prior, both plots showed that $TS > DEG > DG$ in isolation and this is also the result of the competition game. However, for the Heavy Tail prior, we see that $DG, DEG > TS$ for low warm start and that, in general, $DEG > DG$ \footnote{The result that $DEG > DG$ for this prior is robust across decision rules as well.}. While the ``absolute" reputation plots imply that $TS > DEG > DG$, we see that the relative reputation plot in Figure \ref{relative_rep} implies that $DEG > DG$ and in Figure \ref{rel_rep_ht_fine} we can see that in the early rounds, $DG$ does relatively better than $TS$ and that, around $t = 100$, $TS$ does relatively better than $DG$. Note that since the median effective end of game is $0$ in \ref{hm_ht} and the agents are using the $HardMax$ decision rule, this means that the relative reputation between the two firms after $k$ samples ends up determining who takes the market in the simulations and this is precisely what the relative reputation plots tell us. Thus, looking at the relative reputation plot at time $k$ gives us a good predictor of which algorithm will take more of the market in the competition game.

Why is it that we see that $TS$ always wins under Needle in Haystack but, for lower warm start, $DG > TS$? We know that for a sufficiently long time horizon, $TS$ leads to a higher reputation than $DG$. However, in our model, firms need to incentivize the agents to select them over their competition. One possible intuition for the results here are that, in the early rounds, $TS$ engages in purposeful exploration whereas $DG$ is myopic. A consequence of exploration in our model is that it may lead to a reputational cost relative to the greedy algorithm since the sub-optimal arms pulled during exploration can harm reputation relative to the greedy algorithm. $TS$ thus incurs a relative reputational cost in the early rounds, but once it has explored enough and gained enough information it is able to recover the relative reputational loss and acquire a larger reputation than the greedy algorithm. A sufficiently large warm start allows $TS$ to explore in the early rounds and recover the reputational loss and thus win the competition game whereas a low warm start is not sufficient for this and thus $DG$ beats $TS$. Why do we only see this on Heavy Tail and not Needle in Haystack? One possible explanation for this is that learning on Heavy Tail is harder (use learning hardness metric here) and learning on Needle In Haystack is easier and thus TS needs more samples in the Heavy Tail instances compared to the Needle In Haystack instances in order to be able to recover the relative reputation loss [TODO: Explain this a bit more and with more formality. Can we have some experimental results that ``prove" this or generalize it?]

@@
